On the marginal value of swimming in woodland caribou

### 2021-04-12 paper/manuscript.Rmd
Update manuscript.Rmd
P3

Movement is costly. Even for adept swimmers like caribou (Rangifer tarandus), which have hollow hair shafts that aid in flotation, the energetic expenditure of swimming for terrestrial mammals is markedly higher than walking or running (Miller and Gunn 1985). The drivers that promote the decision to swim for terrestrial mammals, like caribou, are often related to the profitability of forage (Miller et al. 1977) and the risks of predation (Jeffery et al. 2007, Jordan et al. 2010). The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to approximately 300 woodland caribou, descendants of 26 individuals introduced from the Island of Newfoundland to Fogo Island between 1964–67 (Bergerud and Mercer 1989). During fieldwork on May 30, 2017, we observed an unmarked adult male caribou swim between two smaller islands (Figure 1) a distance of 470 m which took approximately 9 minutes. Given that swimming is energetically expensive, we ask how often do caribou swim between islands, how long do they occupy islands, and ultimately what motivates swimming? Here, we frame swimming and island residency as a model of patch use as described by the marginal value theorem (Charnov 1976) within the context of the forage limitation hypothesis. For caribou, the forage limitation hypothesis is the constraints on food availability associated with changes in abundance through time (Klein 1991, Schaefer et al. 2016). Here, we integrate our understanding of forage limitation in caribou populations by using island residency as a proxy for time in patch and island size as a proxy for energy intake in a patch (Figure 2).

@@

>>>>

Movement is costly. Even for adept swimmers like caribou (*Rangifer tarandus*), which have hollow hair shafts that aid in flotation, the energetic expenditure of swimming for terrestrial mammals is markedly higher than walking or running (Miller and Gunn 1985). The drivers that promote the decision to swim for terrestrial mammals, like caribou, are often related to the profitability of forage (Miller et al. 1977) and the risks of predation (Jeffery et al. 2007, Jordan et al. 2010). The Fogo Island archipelago, off the coast of Newfoundland, Canada, is home to approximately 300 woodland caribou, descendants of 26 individuals introduced from the Island of Newfoundland to Fogo Island between 1964–67 (Bergerud and Mercer 1989). During fieldwork on May 30, 2017, we observed an unmarked adult male caribou swim between two smaller islands (Figure 1) a distance of 470 m which took approximately 9 minutes. Given that swimming is energetically expensive, we ask how often do caribou swim between islands, how long do they occupy islands, and ultimately what motivates swimming? Here, we frame swimming and island residency as a model of patch use as described by the marginal value theorem (Charnov 1976) within the context of the forage limitation hypothesis. For caribou, the forage limitation hypothesis is the constraints on food availability associated with changes in abundance through time (Klein 1991, Schaefer et al. 2016). Here, we integrate our understanding of forage limitation in caribou populations by using island residency as a proxy for time in patch and island size as a proxy for energy intake in a patch (Figure 2).

@@

Our focus on forage is due to the limited presence of predators in our system. Throughout their range,
wolves (*Canis lupus*) are the primary predator of caribou, although there are no wolves on the Island
of Newfoundland, predation on adult caribou by coyotes (*Canis latrans*) does occur (Lewis and Mahoney
2014). Within the Fogo Island archipelago, coyotes exist at low density (Huang et al. 2021) and the
probability of occurrence decreases on the smaller islands. The low density of predators throughout
the archipelago is therefore an unlikely driver of swimming in our system, though predation may stillinfluence swimming behaviour for caribou elsewhere in their range (e.g. Jeffery et al. 2007).
